Internal physiology and external cues drive changes in reproductive axis of parenting pigeons

### 2020-03-06 docs/manuscript/elife-v1.tex
done with pub for now

Reproduction is a critical period in an organism's life cycle. Parents
must balance their own needs with the energetic costs and
time-constraints required to successfully reproduce and rear their
offspring to independence(1). A complex series of evolutionarily
conserved physiological processes underlie behaviors typically
associated with reproduction in vertebrates, including courtship,
nesting, and parental care. These processes are primarily mediated by
the reproductive axis, which consists of the hypothalamus in the brain,
the pituitary, and the gonads (i.e., the HPG). Our understanding of the
physiological changes that adults navigate as they transition through
reproductive states are primarily limited to a number of circulating
hormones and a small set of well-known genes (cite). We have yet to come
close to conceptualizing the larger symphony of reproductive events that
occur to yield reproductive and parental care behaviors.

@@

@@

>>>>

Reproduction is a critical period in an organism's life cycle. Parents
must balance their own needs with the energetic costs and
time-constraints required to successfully reproduce and rear their
offspring to independence (cite). A complex series of evolutionarily
conserved physiological processes underlie behaviors typically
associated with reproduction in vertebrates, including courtship,
nesting, and parental care. These processes are primarily mediated by
the reproductive axis, which consists of the hypothalamus in the brain,
the pituitary, and the gonads (i.e., the HPG). Our understanding of the
physiological changes that adults navigate as they transition through
reproductive states are primarily limited to a number of circulating
hormones and a small set of well-known genes (cite). We have yet to come
close to conceptualizing the larger symphony of reproductive events that
occur to yield reproductive and parental care behaviors.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{../../figures/fig1-1}
  \caption{Broad patterns of variation in hypothalamic, pituitary, and gonadal gene expression in male and female rock doves across reproductive and parental transitions. a) Experimental design. b) t-SNE plot highlighting tissue differences in gene expression. Each dot represents 1 sample and is colored by the source tissue (hypothalamus: orange, pituitary: teal, gonads: purple). c) t-SNE plot highlighting sex differences in tissue and sex, as shown in light grey (females) and dark grey (males). t-SNE plots highlight treatment effects in females (d), males (e), the hypothalamus (h), pituitary (i), and gonads (j), with samples colored by parental stage (control: red or , nest-building: orangse, lay: yellow; inc.d3: light green, inc.d9: green, inc.d17: dark green, hatch: dark blue, n5: blue, and n9: light blue.}
  \label{figure:fig1}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{../../figures/fig4-1}
  \caption{Another figure}
  \label{figure:fig4}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{../../figures/fig5-1}
  \caption{Another another figure}
  \label{figure:fig5}
\end{figure}

@@

Results in context to literature.

Significance.

Limitations.

Data accessibility and reproducibility are key challenges for today's
researchers. Using Binder and Shiny, we have provides a means for
scientists, physicians, and the general public to explore the data,
code, and results from this project at
\url{https://raynamharris.shinyapps.io/musicalgenes/}. Like Hipposeq
(Cembrowski et al elife 2016, 2018), this tools allows quick and easy
data exploration of RNAseq data from mutiple cell types that relate to
brain function.

\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{../../figures/supplfig3}
  \caption{R shiny, image 1}
  \label{figure:supplfig3}
\end{figure}
